---
title: "Competition description"
author: "Oliver Hinder"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The files in the course content repo:

  i. Assignment/competition/competition-data.csv
  iii. Assignment/competition/competition-test-x-values.csv
  
are the training set, and test set respectively.
The columns of this dataset are `X1`, ..., `X21`, `outcome`. 
The goal is to predict `outcome` using `X1`, ..., `X21`.
You should use `competition-data.csv` to build your model.
You should submit a csv file called `competition-test-outcome.csv` 
containing a single column of
predictions for the outcome variable
for the `competition-test-x-values.csv` dataset.
Only the instructor has access to the corresponding outcome values
which will be used to evaluate your performance.

You should create a git repo with your code that produces
`competition-test-outcome.csv` and add me as a collaborator. 
If there are multiple files then there should be a readme
giving a brief summary.
Please note that I will look at the git history to ensure
that all team members made meaningful contributions 
to the repo.

Hint: if your team are novice git users then it may be best to work
on separate files in the repo to avoid conflicts.

The assignment will be scored on two criteria:

1. 
Quality of the code in the git (3 points). 
    a. Is the code readable? 
    b. Were sensible models tried? 
    c. Does the revision history contain
    descriptive changes?

2. 
Quality of the predictions on the **test set** in terms of root mean squared
error (RMSE). The scores will be calculated as follows

    i. RSME < 12 = 3 points
    ii. RSME < 8 = 4 points
    iii. RSME < 5 = 5 points
    iv. RSME < 3 = 6 points
    v. RMSE < 2.75 = 6.5 points
    vi. RSME < 2.5 = 7 points
    
Note that you will not know your exact score for this question until the 
assignments are graded. However, you should be able to use the validation
set to get a good estimate.
```{r}
library(tidyverse)
library(caret)
library(modelr)
library(e1071)

```
```{r}
cwd <- dirname(rstudioapi::getSourceEditorContext()$path)
data <- read.csv(
  paste(cwd,"/competition-data.csv",sep=""),
  stringsAsFactors = FALSE
)
```

#Dealing with missing values
```{r}
total_na <- function(x) sum(is.na(x))
CPM_na <- sort(data %>% summarise(across(everything(), total_na)),decreasing=TRUE)
CPM_na[,1:6]
```

#Transformations to the predictor distribution
```{r}
skewness(data$outcome)
```

```{r}
data <- data %>% filter(outcome >0)
skewness(data$outcome)
```

```{r}
predictors_only <- select(data, -outcome)
df_predictors_only <- data.frame(predictors_only)
preprocessing_fit <- preProcess(
df_predictors_only,
method = c("BoxCox", "center", "scale"))
preprocessing_fit
```


```{r}
transformed_predictors <- predict(preprocessing_fit, df_predictors_only)
transformed_predictors$outcome = data$outcome
```

#Detecting and handling outliers
```{r}
ggplot(transformed_predictors, aes(y= outcome)) + geom_boxplot()
```

#Adding predictors
```{r}
sapply(transformed_predictors, class)
```

#Data reduction
```{r}
remove_these_predictors <- nearZeroVar(transformed_predictors)
preprocessed <- transformed_predictors[,-remove_these_predictors]
ncol(preprocessed)
```
# preprocessing
```{r}
predictors_only <- data.frame(preprocessed %>% select(-outcome))
predictors_only_correlation_matrix <- cor(predictors_only)
cutoff <- 0.95
names_of_predictors_to_remove <- findCorrelation(
predictors_only_correlation_matrix, names=TRUE, cutoff=cutoff)
removed_predictors <- preprocessed %>%
select(-all_of(names_of_predictors_to_remove))
ncol(removed_predictors)

# folding
trainData <- removed_predictors
folds <- createFolds(trainData$outcome, k = 10, returnTrain = TRUE)
ctrl <- trainControl(method = "cv", index = folds)
```

```{r}
# Non-linear models

# Support Vector Machines (SVM)
set.seed(123)
svm_fit <- train(outcome ~ ., data = trainData, method = "svmRadial", trControl =
                   ctrl)

# Neural Nets
set.seed(123)
nnet_fit <- train(outcome ~ ., data = trainData, method = "nnet", trControl =
                    ctrl)

# Multivariate Adaptive Regression Splines (MARS)
set.seed(123)
mars_fit <- train(outcome ~ ., data = trainData, method = "earth", trControl =
                    ctrl)

# K-nearest
set.seed(123)
knn_fit = train(outcome~ ., data = trainData, method = "knn", trControl =
                  ctrl, preProc = c("center", "scale"))

Non_linear_models <- list(min(svm_fit$results$RMSE),
                          min(nnet_fit$results$RMSE),
                          min(mars_fit$results$RMSE),
                          min(knn_fit$results$RMSE))
Non_linear_models
```

```{r}
# Trees

# Regression tree
set.seed(123)
tree_fit = train(outcome ~ ., data = trainData, method = "rpart", trControl = ctrl)
summary(tree_fit$finalModel)

# Random forests
set.seed(123)
mtryGrid <- data.frame()
rf_fit = train(outcome ~ ., data = trainData, method = "rf", ntree=50, importance = TRUE)
summary(rf_fit$finalModel)

Trees_models <- list(min(tree_fit$results$RMSE),
                     min(rf_fit$results$RMSE))
Trees_models
rf_fit$results
```

```{r}
# Linear Regression

set.seed(100)
lmTune <- train(outcome ~ .,data = trainData,
method = "lm",
trControl = ctrl)

#Ridge

set.seed(100)
ridgeGrid <- expand.grid(lambda = seq(0, .1, length = 10))
ridgeTune <- train(outcome ~ .,data = trainData,
method = "ridge",
tuneGrid = ridgeGrid,
trControl = ctrl,
preProc = c("center", "scale")
)

#Lasso

set.seed(100)
lassoGrid <- expand.grid(lambda = c(0),
                        fraction = seq(.1, 1, length = 15))
  lassoTune <- train(outcome ~ .,data = trainData,
  method = "enet",
  tuneGrid = lassoGrid,
  trControl = ctrl,
  preProc = c("center", "scale")
)

# Elastic net

set.seed(100)
enetGrid <- expand.grid(lambda = seq(0.0, 0.02, length = 5),
                          fraction = seq(.1, 1, length = 5))
  enetTune <- train(outcome ~ .,data = trainData,
  method = "enet",
  tuneGrid = enetGrid,
  trControl = ctrl,
  preProc = c("center", "scale")
)
Regression_models <- list(min(lmTune$results$RMSE),
                          min(ridgeTune$results$RMSE),
                          min(lassoTune$results$RMSE),
                          min(enetTune$results$RMSE))
Regression_models
```

# using the model and doing the output
```{r}
# read the input value
cwd <- dirname(rstudioapi::getSourceEditorContext()$path)
data2 <- read.csv(
  paste(cwd,"/competition-test-x-values.csv",sep=""),
  stringsAsFactors = FALSE
)
```

```{r}
# do the prediction with random forest
predictions <- predict(rf_fit, data2)
data2$outcome <- predictions
```

```{r}
write.csv(data2, paste(cwd,"/results.csv",sep=""), row.names = FALSE)
```
